---
title: "Reproducible Research: Peer Assessment 1"
author: "Taehee"
date: "June 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Course Project 1
This assigment is course project 1 for Reproducible research.

This assignment makes use of data from a personal activity monitoring device.

## Loading and preprocessing the data
```{r loading data}
data<-read.csv("activity.csv")
#head(data)
#steps<-data$steps
#mean(is.na(steps))
data$date <- as.Date(as.character(data$date),"%Y-%m-%d")
```

## Histogram of the total number of steps taken each day
```{r higtogram}
#plot(data$date,data$steps,pch=20)
#Totstep <- with(data, tapply(steps, date, sum, na.rm = TRUE))
activity_total<-aggregate(steps ~ date,data,sum, na.rm = TRUE)
hist(activity_total$steps,xlab="total number of steps taken each day")
```

## Mean and median number of steps taken each day
```{r mean and median}
activity_mean<-aggregate(steps ~ date,data,mean, na.rm = TRUE)
print(activity_mean)
activity_median<-aggregate(steps ~ date,data,median, na.rm = TRUE)
print(activity_median)

```
## Time series plot of the average number of steps taken
```{r time series plot}
activity_interval<-aggregate(steps ~ interval,data,mean,na.rm=TRUE)
with(activity_interval,plot(interval,steps,type="l",ylab="average daily step"))
```

5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
print(activity_interval[which.max(activity_interval$steps),1])
```


## Imputing missing values
the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
print(sum(is.na(data)))
```

create new data set
```{r}
data1<-data
```


filling in all of the missing values with the mean for that 5-minute interval
```{r}
for (i in 1:dim(data1)[1]){
  if (is.na(data1$steps[i])) {
    data1$steps[i]=activity_interval$steps[which(activity_interval$interval==data1$interval[i])]}
}
```


histogram of the total number of steps taken each day
```{r}
activity_total1<-aggregate(steps ~ date,data,sum, na.rm = TRUE)
hist(activity_total1$steps,xlab="total number of steps taken each day after imputing missing values")
```

mean and median total number of steps taken per day
```{r}
activity_mean1<-aggregate(steps ~ date,data1,mean, na.rm = TRUE)
print(activity_mean1)
activity_median1<-aggregate(steps ~ date,data1,median, na.rm = TRUE)
print(activity_median1)
```


the mean and median total number of steps taken per day after Imputing missing values differ from the estimates from the first part of the assignment. 

 
## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend"
```{r weekday}
data1$weekday<-weekdays(data1$date)
data1$weekend[data1$weekday %in% c("Saturday","Sunday")]<-"weekend"
data1$weekend[data1$weekday %in% c("Monday","Tuesday","Wednesday","Thursday","Friday")]<-"weekday"
data1$weekend<-factor(data1$weekend)
```
Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days 
```{r}
library(ggplot2)
g<-ggplot(data1,aes(interval,steps))
g+geom_line()+facet_grid(weekend~.)+ylab("Number of steps")
```





